\bf Title here

### 2022-11-18 project/PaperTemplate/paper.Rmd
added BC and BCa

# Exposition

The portion of the paper where you describe the new technique as though you were teaching it to a classmate (with whatever name you want to give it - Literature, Background, or Exposition) will be based on your annotated bibliography. You can add sources to what you already have in the annotated bibliography (especially if my feedback says you need to!). To be sure that you are on track with the exposition, a full draft of this portion of the paper is due before Thanksgiving break. This will also help with incorporating proper citation early in the writing process, as you’ll have a bibliography that we can implement for this portion of the paper at a minimum.

@@

Bootstrapping is a statistical method of resampling that allows the estimation of a test statistic from an unknown distribution. In particular, bootstrapping is a computational heavy method which is useful for many different situations.

First, we will introduce the non-parametric bootstrap. Suppose we have a random sample $X = (x_1,x_2,\dots,x_n)$ from our unknown distribution, $F$ and a statistic of interest $\hat{\theta} = \hat{\theta}(X)$. Ideally, the desired test statistic could be found by repeatedly sampling new reproductions of $X$ from $F$. However $F$ is unknown so this is not possible. The non-parametric bootstrap creates an estimate $\hat{F}$ from $F$ using our sample $X$ without making any parametric assumptions (such as distribution). Therefore, the bootstrap sample could be represented as $X^* = (x^*_1, x^*_2, \dots, x^*_n)$ where each $x^*_i$ is sampled randomly with equal probability and with replacement from $\{x_1,x_2,\dots,x_n\}$. From this bootstrap sample, a bootstrap replication of the test statistic can be computed using $\hat{\theta}^* = \hat{\theta}(X^*)$. A large number, $B$, of bootstrap samples are drawn independently and the corresponding bootstrap replication of the test statistic is calculated. $$\hat{\theta}^{*b} = \hat{\theta}(X^{*b}) \text{ for } b = 1,2, \dots, B$$. The bootstrap estimate of the test statistic is the empirical value of the test statistic from all of the $\hat{\theta}^{*b}$ replications. As $B$ increases, $\hat{F}$ approaches $F$ which means that the tests statistic of interest approaches its true value as well.

@@

Confidence intervals are tools that are used to estimate a parameter. Specifically, a confidence interval gives a range in which the true value of the parameter may lie. An $\alpha$-level standard confidence interval is given by $$\hat{\theta} \pm z_{\alpha}\hat{\sigma}$$ where $\hat{\theta}$ is a point estimate of the parameter of interest $\theta$, $\hat{\sigma}$ is the estimate of the standard deviation of $\hat{\theta}$ and $z_{\alpha}$ is the $(100 *\alpha)$th percentile of the normal deviation. We say that the confidence interval constructed in this manner has a chance of capturing the true parameter with a probability of $\alpha$.

The standard confidence interval is built based on the assumption that the distribution from which we are sampling is Normal. This means that for an unknown distribution, the standard confidence interval could present an incorrect range. However, the same process can be used with bootstrap sampling to form the bootstrap percentile method. This means that an approximate bootstrap confidence interval will be created in the same automatic way that the standard confidence interval was created.

The percentile method interval is defined as the interval between the $100 * \alpha$ and the $100(1 - \alpha)$ percentiles of the bootstrap distribution of $\hat{\theta}$. That is, the $(1 - 2\alpha)$ coverage interval can be defined as $$[\hat{\theta}^*_\alpha,\hat{\theta}^*_{1-\alpha}].$$ In the case that the bootstrap distribution of $\hat{\theta}^* \sim N(\hat{\theta}, \hat{\sigma}^2)$, the corresponding percentile interval would be equivalent to the standard interval. However, this is not usually the case. When the bootstrap distribution is non-normal, we can suppose that there exists, for all $\theta$, $$\hat{\phi} \sim N(\phi, \tau^2),$$ for some monotone transformation $\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)$, and $\tau$ is a constant. In other words, this transformation perfectly normalizes the distribution of $\hat{\theta}$. This transformation invariant can be applied to the bootstrap replications such that $$\hat{\phi}^{*b} = g\left( \hat{\theta}^{*b}\right ) \text{ for } b = 1,2,\dots, B.$$ The corresponding percentiles of the distribution transform similarly, $\hat{\phi}^*_\alpha = g \left ( \hat{\theta}^*_\alpha \right )$. Or we can say that the $(1 - 2\alpha)$ percentile interval is $\hat{\phi} \pm \tau z_\alpha$ which can also be represented as  $[\hat{\phi}^*_\alpha,\hat{\phi}^*_{1-\alpha}]$. This means that the interval on the $\theta$ scale can be defined as $$\hat{\theta}^*_\alpha = g^{-1}(\hat{\phi} \pm \tau z_\alpha).$$ Or $\left [ g^{-1}(\hat{\phi} \pm \tau z_{1-\alpha}), g^{-1}(\hat{\phi} \pm \tau z_\alpha) \right ]$. Therefore, the percentile method produces a correct interval for $\phi$ and due to the transformation invariance, also produces a correct percentile interval for $\theta$. This method assumes the existence of some monotone normalizing mapping $\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)$ and relies on that to create a correct interval. Since the process is automatic, we do not need to know the transformation itself, only that it exists. However, in some cases, no monotone normalizing mapping will exist.

>>>>

what is exposition
The portion of the paper where you describe the new technique as though you were teaching it to a classmate (with whatever name you want to give it - Literature, Background, or Exposition) will be based on your annotated bibliography. You can add sources to what you already have in the annotated bibliography (especially if my feedback says you need to!). To be sure that you are on track with the exposition, a full draft of this portion of the paper is due before Thanksgiving break. This will also help with incorporating proper citation early in the writing process, as you’ll have a bibliography that we can implement for this portion of the paper at a minimum.

@@

\newpage
# Exposition
Bootstrapping is a statistical method of resampling that allows the estimation of a test statistic from an unknown distribution. In particular, bootstrapping is a computational heavy method which is useful for many different situations.

First, we will introduce the non-parametric bootstrap. Suppose we have a random sample $X = (x_1,x_2,\dots,x_n)$ from our unknown distribution, $F$ and a statistic of interest $\hat{\theta} = \hat{\theta}(X)$. Ideally, the desired test statistic could be found by repeatedly sampling new reproductions of $X$ from $F$. However $F$ is unknown so this is not possible. The non-parametric bootstrap creates an estimate $\hat{F}$ from $F$ using our sample $X$ without making any parametric assumptions (such as distribution). Therefore, the bootstrap sample could be represented as $X^* = (x^*_1, x^*_2, \dots, x^*_n)$ where each $x^*_i$ is sampled randomly with equal probability and with replacement from $\{x_1,x_2,\dots,x_n\}$. From this bootstrap sample, a bootstrap replication of the test statistic can be computed using $\hat{\theta}^* = \hat{\theta}(X^*)$. A large number, $B$, of bootstrap samples are drawn independently and the corresponding bootstrap replication of the test statistic is calculated. $$\hat{\theta}^{*b} = \hat{\theta}(X^{*b}) \text{ for } b = 1,2, \dots, B.$$ The bootstrap estimate of the test statistic is the empirical value of the test statistic from all of the $\hat{\theta}^{*b}$ replications. As $B$ increases, $\hat{F}$ approaches $F$ which means that the tests statistic of interest approaches its true value as well.

@@

Confidence intervals are tools that are used to estimate a parameter. Specifically, a confidence interval gives a range in which the true value of the parameter may lie. An $\alpha$-level standard confidence interval is given by $$\hat{\theta}_S[\alpha] = \hat{\theta} \pm z_{\alpha}\hat{\sigma}$$ where $\hat{\theta}$ is a point estimate of the parameter of interest $\theta$, $\hat{\sigma}$ is the estimate of the standard deviation of $\hat{\theta}$ and $z_{\alpha}$ is the $(100 *\alpha)$th percentile of the normal deviation. We say that the confidence interval constructed in this manner has a chance of capturing the true parameter with a probability of $\alpha$.

The standard confidence interval is built based on the assumption that the distribution from which we are sampling is Normal. This means that for an unknown distribution, the standard confidence interval could present an incorrect range. However, the same process can be used with bootstrap sampling to form the bootstrap percentile method. This means that an approximate bootstrap confidence interval will be created in the same automatic way that the standard confidence interval was created. For bootstrapped confidence intervals, the number of bootstrap replications $B$ must be large (around 2000) due to the nature of confidence intervals requiring greater accuracy.

The percentile method interval is defined as the interval between the $100 * \alpha$ and the $100(1 - \alpha)$ percentiles of the bootstrap distribution of $\hat{\theta}$. That is, the $(1 - 2\alpha)$ coverage interval can be defined as $[\hat{\theta}^*_\alpha,\hat{\theta}^*_{1-\alpha}].$ To go further, we can say define $\hat{G}(t)$ as the bootstrap cdf, or the proportion of bootstrap samples less than $t$. $$\hat{G}(t) = \frac{\#\{\hat{\theta^{*b} \leq t}\}}{B}$$ Thus the $\alpha$th percentile point of the distribution is given by $$\hat{\theta}_p[\alpha] = \hat{\theta^*_\alpha} = \hat{G}^{-1}(\alpha).$$

It follows that the percentile interval can be represented as $$\left [ \hat{G}^{-1}(\alpha),\hat{G}^{-1}(1-\alpha) \right ].$$

In the case that the bootstrap distribution of $\hat{\theta}^* \sim N(\hat{\theta}, \hat{\sigma}^2)$, the corresponding percentile interval would be equivalent to the standard interval. However, this is not usually the case. When the bootstrap distribution is non-normal, we can suppose that there exists, for all $\theta$, $$\hat{\phi} \sim N(\phi, \tau^2),$$ for some monotone transformation $\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)$, and $\tau$ is a constant. In other words, this transformation perfectly normalizes the distribution of $\hat{\theta}$. This transformation invariant can be applied to the bootstrap replications such that $$\hat{\phi}^{*b} = g\left( \hat{\theta}^{*b}\right ) \text{ for } b = 1,2,\dots, B.$$ The corresponding percentiles of the distribution transform similarly, $\hat{\phi}^*_\alpha = g \left ( \hat{\theta}^*_\alpha \right )$. Or we can say that the $(1 - 2\alpha)$ percentile interval is $\hat{\phi} \pm \tau z_\alpha$ which can also be represented as  $[\hat{\phi}^*_\alpha,\hat{\phi}^*_{1-\alpha}]$. This means that the interval on the $\theta$ scale can be defined as $$\hat{\theta}^*_\alpha = g^{-1}(\hat{\phi} \pm \tau z_\alpha).$$ Or $\left [ g^{-1}(\hat{\phi} \pm \tau z_{1-\alpha}), g^{-1}(\hat{\phi} \pm \tau z_\alpha) \right ]$. Therefore, the percentile method produces a correct interval for $\phi$ and due to the transformation invariance, also produces a correct percentile interval for $\theta$. This method assumes the existence of some monotone normalizing mapping $\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)$ and relies on that to create a correct interval. Since the process is automatic, we do not need to know the transformation itself, only that it exists. However, in some cases, no monotone normalizing mapping will exist.

The next method we will be looking at is the bias-corrected percentile method (BC method) which is an improvement upon the previous percentile method because we now take into account the possibility of bias. It can be shown (SHOULD I SHOW LATER?) that $\hat{\theta}$ is biased upwards relative to $\theta$ which means that the confidence intervals should be adjusted downwards. From our simulated bootstrap replications $\hat{\theta}^{*1}, \hat{\theta}^{*2}, \dots ,\hat{\theta}^{*B}$, define $p_0 = = \frac{\#\{\hat{\theta^{*b} \leq \theta}\}}{B}$, and define the bias-correction value $z_0 = \Phi^{-1}(p_0)$ where $\Phi^{-1}$ is the inverse function of the standard normal cdf. Thus we define a transformation $\hat{\phi} = g(\hat{\theta)}, \phi = g(\hat{\theta)}$ such that for any $\theta$, $\hat{\phi} \sim N(\phi - z_0\tau, \tau^2)$, with $z_0$ and $\tau$ constants. This means that we can say the bias corrected method has an $\alpha$-level endpoint can be represented as

$$\hat{\theta}_{BC}[\alpha] = \hat{G}^{-1} \left [ \Phi \left ( 2z_0 + z_\alpha\right ) \right ].$$ If $\hat{G} = 0.50$, then half of the bootstrap distribution is less than $\hat{\theta}$ and our bias-correction value $z_0 = 0$. In this case, the confidence interval produced by BC would be the same interval produced by the percentile method.

A further modification upon the BC interval is the bias corrected and accelerated method (BCa). For this method, we do assume the the standard error, $\tau$ to be constant as we did in the BC interval. Rather, we assume the existence of a monotone transformation $\hat{\phi} = g(\hat{\theta)}, \phi = g(\hat{\theta)}$ such that for any $\theta$, $\hat{\phi} \sim N(\phi - z_0\tau_\phi, \tau_\phi^2)$ where $\tau_\phi = 1+ a\phi$. The $a$ in known as the acceleration value and is a constant that describes how the standard deviation of $\hat{\phi}$ varies with $\phi$. In other words, $a$ is proportional to the skewness of the bootstrap distribution. For example, for one-parameter exponential families, $a=z_0$, however, there are many different algorithms to compute and estimate $a$. Now, our $\alpha$-level endpoint from BCa is $$\hat{\theta}_{BCa}[\alpha] = \hat{G}^{-1} \left [ \Phi \left ( z_0 + \frac{z_0+z_\alpha}{1-a(z_0+z_a)} \right ) \right ].$$ If $a = 0$, then $\hat{\theta}_{BCa}[\alpha] = \hat{\theta}_{BC}[\alpha].$ When calculating a BCa interval, the acceleration value $a$ is not a function of the bootstrap distribution and must be calculated separate, however the process is algorithmic and can be calculated without too much work. Each of the three previous methods (percentile, BC, and BCa) all build upon each other and have less restrictive assumptions, however computation increases as we loosen assumptions.

### 2022-11-18 project/PaperTemplate/paper.tex
added BC and BCa

The standard confidence interval is built based on the assumption that
the distribution from which we are sampling is Normal. This means that
for an unknown distribution, the standard confidence interval could
present an incorrect range. However, the same process can be used with
bootstrap sampling to form the bootstrap percentile method. This means
that an approximate bootstrap confidence interval will be created in the
same automatic way that the standard confidence interval was created.

The percentile method interval is defined as the interval between the
\(100 * \alpha\) and the \(100(1 - \alpha)\) percentiles of the
bootstrap distribution of \(\hat{\theta}\). That is, the
\((1 - 2\alpha)\) coverage interval can be defined as
\[[\hat{\theta}^*_\alpha,\hat{\theta}^*_{1-\alpha}].\] In the case that
the bootstrap distribution of
\(\hat{\theta}^* \sim N(\hat{\theta}, \hat{\sigma}^2)\), the
corresponding percentile interval would be equivalent to the standard
interval. However, this is not usually the case. When the bootstrap
distribution is non-normal, we can suppose that there exists, for all
\(\theta\), \[\hat{\phi} \sim N(\phi, \tau^2),\] for some monotone
transformation \(\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)\), and
\(\tau\) is a constant. In other words, this transformation perfectly
normalizes the distribution of \(\hat{\theta}\). This transformation
invariant can be applied to the bootstrap replications such that
\[\hat{\phi}^{*b} = g\left( \hat{\theta}^{*b}\right ) \text{ for } b = 1,2,\dots, B.\]
The corresponding percentiles of the distribution transform similarly,
\(\hat{\phi}^*_\alpha = g \left ( \hat{\theta}^*_\alpha \right )\). Or
we can say that the \((1 - 2\alpha)\) percentile interval is
\(\hat{\phi} \pm \tau z_\alpha\) which can also be represented as
\([\hat{\phi}^*_\alpha,\hat{\phi}^*_{1-\alpha}]\). This means that the
interval on the \(\theta\) scale can be defined as
\[\hat{\theta}^*_\alpha = g^{-1}(\hat{\phi} \pm \tau z_\alpha).\] Or
\(\left [ g^{-1}(\hat{\phi} \pm \tau z_{1-\alpha}), g^{-1}(\hat{\phi} \pm \tau z_\alpha) \right ]\).
Therefore, the percentile method produces a correct interval for
\(\phi\) and due to the transformation invariance, also produces a
correct percentile interval for \(\theta\). This method assumes the
existence of some monotone normalizing mapping
\(\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)\) and relies on that to
create a correct interval. Since the process is automatic, we do not
need to know the transformation itself, only that it exists. However, in
some cases, no monotone normalizing mapping will exist.

>>>>

The standard confidence interval is built based on the assumption that
the distribution from which we are sampling is Normal. This means that
for an unknown distribution, the standard confidence interval could
present an incorrect range. However, the same process can be used with
bootstrap sampling to form the bootstrap percentile method. This means
that an approximate bootstrap confidence interval will be created in the
same automatic way that the standard confidence interval was created.
For bootstrapped confidence intervals, the number of bootstrap
replications \(B\) must be large (around 2000) due to the nature of
confidence intervals requiring greater accuracy.

The percentile method interval is defined as the interval between the
\(100 * \alpha\) and the \(100(1 - \alpha)\) percentiles of the
bootstrap distribution of \(\hat{\theta}\). That is, the
\((1 - 2\alpha)\) coverage interval can be defined as
\([\hat{\theta}^*_\alpha,\hat{\theta}^*_{1-\alpha}].\) To go further, we
can say define \(\hat{G}(t)\) as the bootstrap cdf, or the proportion of
bootstrap samples less than \(t\).
\[\hat{G}(t) = \frac{\#\{\hat{\theta^{*b} \leq t}\}}{B}\] Thus the
\(\alpha\)th percentile point of the distribution is given by
\[\hat{\theta}_p[\alpha] = \hat{\theta^*_\alpha} = \hat{G}^{-1}(\alpha).\]

It follows that the percentile interval can be represented as
\[\left [ \hat{G}^{-1}(\alpha),\hat{G}^{-1}(1-\alpha) \right ].\]

In the case that the bootstrap distribution of
\(\hat{\theta}^* \sim N(\hat{\theta}, \hat{\sigma}^2)\), the
corresponding percentile interval would be equivalent to the standard
interval. However, this is not usually the case. When the bootstrap
distribution is non-normal, we can suppose that there exists, for all
\(\theta\), \[\hat{\phi} \sim N(\phi, \tau^2),\] for some monotone
transformation \(\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)\), and
\(\tau\) is a constant. In other words, this transformation perfectly
normalizes the distribution of \(\hat{\theta}\). This transformation
invariant can be applied to the bootstrap replications such that
\[\hat{\phi}^{*b} = g\left( \hat{\theta}^{*b}\right ) \text{ for } b = 1,2,\dots, B.\]
The corresponding percentiles of the distribution transform similarly,
\(\hat{\phi}^*_\alpha = g \left ( \hat{\theta}^*_\alpha \right )\). Or
we can say that the \((1 - 2\alpha)\) percentile interval is
\(\hat{\phi} \pm \tau z_\alpha\) which can also be represented as
\([\hat{\phi}^*_\alpha,\hat{\phi}^*_{1-\alpha}]\). This means that the
interval on the \(\theta\) scale can be defined as
\[\hat{\theta}^*_\alpha = g^{-1}(\hat{\phi} \pm \tau z_\alpha).\] Or
\(\left [ g^{-1}(\hat{\phi} \pm \tau z_{1-\alpha}), g^{-1}(\hat{\phi} \pm \tau z_\alpha) \right ]\).
Therefore, the percentile method produces a correct interval for
\(\phi\) and due to the transformation invariance, also produces a
correct percentile interval for \(\theta\). This method assumes the
existence of some monotone normalizing mapping
\(\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)\) and relies on that to
create a correct interval. Since the process is automatic, we do not
need to know the transformation itself, only that it exists. However, in
some cases, no monotone normalizing mapping will exist.

The next method we will be looking at is the bias-corrected percentile
method (BC method) which is an improvement upon the previous percentile
method because we now take into account the possibility of bias. It can
be shown (SHOULD I SHOW LATER?) that \(\hat{\theta}\) is biased upwards
relative to \(\theta\) which means that the confidence intervals should
be adjusted downwards. From our simulated bootstrap replications
\(\hat{\theta}^{*1}, \hat{\theta}^{*2}, \dots ,\hat{\theta}^{*B}\),
define \(p_0 = = \frac{\#\{\hat{\theta^{*b} \leq \theta}\}}{B}\), and
define the bias-correction value \(z_0 = \Phi^{-1}(p_0)\) where
\(\Phi^{-1}\) is the inverse function of the standard normal cdf. Thus
we define a transformation
\(\hat{\phi} = g(\hat{\theta)}, \phi = g(\hat{\theta)}\) such that for
any \(\theta\), \(\hat{\phi} \sim N(\phi - z_0\tau, \tau^2)\), with
\(z_0\) and \(\tau\) constants. This means that we can say the bias
corrected method has an \(\alpha\)-level endpoint can be represented as

\[\hat{\theta}_{BC}[\alpha] = \hat{G}^{-1} \left [ \Phi \left ( 2z_0 + z_\alpha\right ) \right ].\]
If \(\hat{G} = 0.50\), then half of the bootstrap distribution is less
than \(\hat{\theta}\) and our bias-correction value \(z_0 = 0\). In this
case, the confidence interval produced by BC would be the same interval
produced by the percentile method.

A further modification upon the BC interval is the bias corrected and
accelerated method (BCa). For this method, we do assume the the standard
error, \(\tau\) to be constant as we did in the BC interval. Rather, we
assume the existence of a monotone transformation
\(\hat{\phi} = g(\hat{\theta)}, \phi = g(\hat{\theta)}\) such that for
any \(\theta\), \(\hat{\phi} \sim N(\phi - z_0\tau_\phi, \tau_\phi^2)\)
where \(\tau_\phi = 1+ a\phi\). The \(a\) in known as the acceleration
value and is a constant that describes how the standard deviation of
\(\hat{\phi}\) varies with \(\phi\). In other words, \(a\) is
proportional to the skewness of the bootstrap distribution. For example,
for one-parameter exponential families, \(a=z_0\), however, there are
many different algorithms to compute and estimate \(a\). Now, our
\(\alpha\)-level endpoint from BCa is
\[\hat{\theta}_{BCa}[\alpha] = \hat{G}^{-1} \left [ \Phi \left ( z_0 + \frac{z_0+z_\alpha}{1-a(z_0+z_a)} \right ) \right ].\]
If \(a = 0\), then
\(\hat{\theta}_{BCa}[\alpha] = \hat{\theta}_{BC}[\alpha].\) When
calculating a BCa interval, the acceleration value \(a\) is not a
function of the bootstrap distribution and must be calculated separate,
however the process is algorithmic and can be calculated without too
much work. Each of the three previous methods (percentile, BC, and BCa)
all build upon each other and have less restrictive assumptions, however
computation increases as we loosen assumptions.
